Smart Spreading Factor Assignment for LoRaWANs

### 2019-05-01 paper.tex
prediction energy figure is added

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{collision.png}
\caption{Collision between nodes close to the gateway.}
\label{fig:collision}
\end{figure}

@@

The adaptive data rate (ADR) algorithm recommended by Semtech Corporation utilizes signal to noise ratio (SNR) of the last 20 transmissions to minimize the transmit power and SF while ensuring successful transmissions \cite{lorawan_adr}. The recommended algorithm uses a predefined SF and corresponding required SNR table. The algorithm considers maximum SNR of the last 20 transmissions. The algorithm increases transmit power or decrease data rate in case of low SNR and the opposite in case of high SNR.

@@

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{collision_solution_single_gw.png}
\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
\label{fig:collision_solution_single_gw}
\end{figure}

@@

\begin{table}
\centering
\caption{EU863-870 ISM band LoRa default channels \cite{lorawan.regional.parameters}.}
\label{table:max_tx_power}
\begin{tabular}{|c|c|c|}
\hline
\textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
      868.1 &   125 &   14 \\ \hline
      868.3 &   125 &   14 \\ \hline
      868.5 &   125 &   14 \\ \hline
\end{tabular}
\end{table}

Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}. In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm. Different channel transmissions are independent from each other. But, in this work we focus on SF orthogonality. Thus, only single channel transmissions are utilized in the simulator.

@@

Where $t_{x}$ is transmission duration of referenced signal. $t_{interf}$ is overlapping duration between referenced signal and interfering signal. Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$. Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}. Duration of a LoRa transmission can be calculated as:

\begin{equation} \label{eq:transmission_duration}
t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
\end{equation}

@@

\subsection{Smart SF Schemes}
PDR plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of gateways is set to 3. Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway. If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

@@

>>>>

\begin{figure}
\centering
\includegraphics[width=0.83\linewidth]{collision.png}
\caption{Collision between nodes close to the gateway.}
\label{fig:collision}
\end{figure}

@@

The adaptive data rate (ADR) algorithm recommended by Semtech Corporation utilizes signal to noise ratio (SNR) of the last 20 transmissions to minimize the transmit power and SF while ensuring successful transmissions \cite{lorawan_adr}. The recommended algorithm uses a predefined SF and corresponding required SNR table. The algorithm increases transmit power or decrease data rate in case of low SNR and the opposite in case of high SNR.

@@

\begin{figure}
\centering
\includegraphics[width=0.83\linewidth]{collision_solution_single_gw.png}
\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
\label{fig:collision_solution_single_gw}
\end{figure}

@@

In the simulator, it is assumed that nodes always select maximum allowed transmit power for European ISM band, which is 14 dBm \cite{lorawan.regional.parameters}. Different channel transmissions are independent from each other. But, in this work we focus on SF orthogonality. Thus, only single channel transmissions are utilized in the simulator.

@@

Where $t_{x}$ is transmission duration of referenced signal. $t_{interf}$ is overlapping duration between referenced signal and interfering signal. Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$. Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.

@@

\subsection{Smart SF Schemes}
PDR and transmit energy consumption plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr} and \ref{fig:prediction_energy}. Randomly generated network radius is set to 5000 meters and number of gateways is set to 3. Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Random SF scheme consumes far more energy than others, therefore it is not feasible for real world deployments. As number of nodes increases, both smart SVM and smart DTC schemes consume slightly more energy than lowest SF scheme, however they give better PDR results in return. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better PDR results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway. If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{prediction_energy_r5000_g3_p0.01_s3600}.png}
\caption{TX energy for lowest and smart SF schemes. (r = 5000 m, GW = 3)}
\label{fig:prediction_energy}
\end{figure}
